HTW Model e3

### 2024-02-24 Model/htw_model_e3.qmd
add prelim e3 modelling

---
title: HTW Model e2
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

#ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
ds <- readRDS(here::here("data/e2_md_02-23-24.rds"))  |> as.data.table()
nbins <- 3

fd <- readRDS(here("data/e2_08-21-23.rds"))
test <- fd |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

@@

file_name <- "e2_n_iter_50_ntry_200_2506"
#file_name <- "n_iter_400_ntry_100_2944"
file_name <- "e2_n_iter_100_ntry_200_3436"

@@

# organize test data predictions
 post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
    generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 50, return_dat="test_data")
    })) |> 
   select(Fit_Method,pp,-data) |>  
   unnest(pp) |>  filter(expMode2=="Test") |> as.data.table()
# 
# saveRDS(post_dat, here("data/model_cache/post_dat.rds"))

@@

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  mutate(error2 = y - pred) |>
  summarise(y = mean(y), pred = mean(pred), error = y - pred, error2=mean(error2)) |> as.data.table()

setorder(post_dat_avg, id, x, rank)
post_dat_l <- melt(post_dat_avg, id.vars = c("id", "condit", "Model", "Fit_Method", "x", "c", "lr", "rank","error"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val")
post_dat_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))]
setorder(post_dat_l, id, Resp)
#rm(post_dat_avg)

@@

@@

```{r fig.height=12,fig.width=11}
#| label: fig-htw-resid-pred
#| column: page-inset-right
#| fig-cap: A) Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets. B) Full posterior predictive distributions vs. observed data from participants.Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. 
#| fig-height: 19
#| fig-width: 12
##| layout: [[45,-5, 45], [100]]
##| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data","Full posterior predictive distributions vs. observed data from participants."]
train_resid <- pd_train |> group_by(id,condit,Model,Fit_Method, Block) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |>
  ggplot(aes(x = Block, y = abs(error), fill=Model)) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, scales="free",ncol=2) +
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  labs(title="Model Residual Errors - Training Stage", y="RMSE", x= "Training Block") +
  theme(legend.title = element_blank(), legend.position="top")

test_resid <-  post_dat |> 
   group_by(id,condit,x,Model,Fit_Method,rank) |>
   summarize(error=mean(abs(y-pred)),n=n()) |>
   group_by(id,condit,x,Model,Fit_Method) |>
   summarize(error=mean(error)) |>
  mutate(vbLab = factor(paste0(x,"-",x+200))) |>
  ggplot(aes(x = vbLab, y = abs(error), fill=Model)) + 
  stat_bar + 
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Model Residual Errors - Testing Stage",y="RMSE", x="Velocity Band") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

group_pred <- post_dat_l |> 
  mutate(vbLab = factor(paste0(x,"-",x+200),levels=levels(testAvg$vb))) |>
  ggplot(aes(x=val,y=vbLab,col=Resp)) + 
  stat_pointinterval(position=position_dodge(.5), alpha=.9) + 
  scale_color_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit, axes = "all",ncol=2,scale="free") +
  labs(title="Posterior Predictions - Testing Stage",y="Velocity Band (lower bound)", x="X Velocity") +
theme(legend.title=element_blank(),axis.text.y = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

@@

c_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=log(c), x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="c parameter") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4))

lr_post <- post_dat_avg %>%
    group_by(id, condit, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=lr, x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.4)) +
    ggh4x::facet_nested_wrap(~Model) + labs(title="learning rate parameter") +
  theme(legend.title = element_blank(), legend.position = "none",plot.title=element_text(hjust=.5))
c_post + lr_post
```

@@

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, Error_difference)) %>%
  ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
  geom_col() +
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")+
  labs(fill="Best Model",x="Mean Model Error Difference (ALM - EXAM)",y="Participant")

@@

#### Subjects with biggest differential favoring ALM
```{r}
#| fig-height: 6
#| fig-width: 12
vAlm <- c(307,331,197); cAlm <- c(372,173,157)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Subjects with biggest differential favoring ALM",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

@@

#### Subjects with biggest differential favoring EXAM
```{r}
#| fig-height: 6
#| fig-width: 12
vAlm <- c(312,334,295); cAlm <- c(132,366,415)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Subjects with biggest differential favoring EXAM",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

>>>>

---
title: HTW Model e3
author: Thomas Gorman
date: "`r Sys.Date()`"
page-layout: full
toc: false
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

@@

#ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
ds <- readRDS(here::here("data/e3_md_02-23-24.rds"))  |> as.data.table()
nbins <- 3

fd <- readRDS(here("data/e3_08-21-23.rds"))
group_ids <- fd |> distinct(id,condit,bandOrder,fb)
test <- fd |> filter(expMode2 == "Test") 
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

@@

#file_name <- "e2_n_iter_50_ntry_200_2506"
#file_name <- "n_iter_400_ntry_100_2944"
#file_name <- "e2_n_iter_100_ntry_200_3436"
file_name <- "e3_n_iter_50_ntry_100_3244"

@@

# organize test data predictions
 post_dat <- nestSbjModelFit |> mutate(pp=furrr::future_pmap(list(id,Model,Fit_Method,data), ~{
    generate_data(..2, ..4 |> mutate(id=..1), ds, num_samples = 30, return_dat="test_data")
    })) |> 
   select(Fit_Method,pp,-data) |>  
   unnest(pp) |>  filter(expMode2=="Test") |> as.data.table()
# 
# saveRDS(post_dat, here("data/model_cache/post_dat.rds"))

@@

post_dat_avg <- post_dat |> group_by(id, condit, Model, Fit_Method, x, c, lr, rank) |> 
  mutate(error2 = y - pred) |>
  summarise(y = mean(y), pred = mean(pred), error = y - pred, error2=mean(error2)) |> as.data.table() |>
  left_join(group_ids, by = join_by(id,condit))

setorder(post_dat_avg, id, x, rank)
post_dat_l <- melt(post_dat_avg, id.vars = c("id", "condit","bandOrder","fb", "Model", "Fit_Method", "x", "c", "lr", "rank","error"),
                   measure.vars = c("pred", "y"), variable.name = "Resp", value.name = "val")
post_dat_l[, Resp := fifelse(Resp == "y", "Observed",
                             fifelse(Model == "ALM", "ALM", "EXAM"))]
setorder(post_dat_l, id, Resp)
#rm(post_dat_avg)

@@

post_dat  |> 
  left_join(group_ids, by = join_by(id,condit)) |>
  group_by(id,condit,bandOrder,Model,Fit_Method,x) |>
  mutate(e2=abs(y-pred)) |>
  summarise(y1=mean(y), pred1=mean(pred)) |>
  group_by(condit,bandOrder,Model,Fit_Method) |>
  summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |>
  # group_by(condit,bandOrder,Model,Fit_Method) |>
  # summarise(mean_error=mean(mean_error)) |>
  round_tibble(1) |> 
  flextable::tabulator(rows=c("Fit_Method","Model"), columns=c("condit","bandOrder"),
                       `ME` = as_paragraph(mean_error)) |> as_flextable()

@@

```{r fig.height=12,fig.width=11}
#| label: fig-htw-resid-pred
#| column: page-inset-right
#| fig-cap: A) Model residuals for each combination of training condition, fit method, and model. Residuals reflect the difference between observed and predicted values. Lower values indicate better model fit. Note that y axes are scaled differently between facets. B) Full posterior predictive distributions vs. observed data from participants.Points represent median values, thicker intervals represent 66% credible intervals and thin intervals represent 95% credible intervals around the median. 
#| fig-height: 32
#| fig-width: 14
##| layout: [[45,-5, 45], [100]]
##| fig-subcap: ["Model Residuals - training data", "Model Residuals - testing data","Full posterior predictive distributions vs. observed data from participants."]
train_resid <- pd_train |> 
  left_join(group_ids, by = join_by(id,condit)) |>
  group_by(id,condit,bandOrder,Model,Fit_Method, Block) |> 
  summarise(y = mean(y), pred = mean(pred), error = y - pred) |>
  ggplot(aes(x = Block, y = abs(error), fill=Model)) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit~bandOrder, scales="free",ncol=2) +
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  labs(title="Model Residual Errors - Training Stage", y="RMSE", x= "Training Block") +
  theme(legend.title = element_blank(), legend.position="top")

test_resid <-  post_dat |> 
  left_join(group_ids, by = join_by(id,condit)) |>
   group_by(id,condit,bandOrder,x,Model,Fit_Method,rank) |>
   summarize(error=mean(abs(y-pred)),n=n()) |>
   group_by(id,condit,bandOrder,x,Model,Fit_Method) |>
   summarize(error=mean(error)) |>
  mutate(vbLab = factor(paste0(x,"-",x+200))) |>
  ggplot(aes(x = vbLab, y = abs(error), fill=Model)) + 
  stat_bar + 
  scale_fill_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit~bandOrder, axes = "all",ncol=2,scale="free") +
  labs(title="Model Residual Errors - Testing Stage",y="RMSE", x="Velocity Band") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

group_pred <- post_dat_l |> 
  mutate(vbLab = factor(paste0(x,"-",x+200),levels=levels(testAvg$vb))) |>
  ggplot(aes(x=val,y=vbLab,col=Resp)) + 
  stat_pointinterval(position=position_dodge(.5), alpha=.9) + 
  scale_color_manual(values=wes_palette("AsteroidCity2"))+
  ggh4x::facet_nested_wrap(rename_fm(Fit_Method)~condit~bandOrder, axes = "all",ncol=2,scale="free") +
  labs(title="Posterior Predictions - Testing Stage",y="Velocity Band (lower bound)", x="X Velocity") +
theme(legend.title=element_blank(),axis.text.y = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

@@

c_post <- post_dat_avg %>%
    group_by(id, condit,bandOrder, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=log(c), x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.2)) +
    ggh4x::facet_nested_wrap(~Model+bandOrder) + labs(title="c parameter") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4))

lr_post <- post_dat_avg %>%
    group_by(id, condit,bandOrder, Model, Fit_Method, rank) %>%
    slice_head(n = 1) |>
    ggplot(aes(y=lr, x = Fit_Method,col=condit)) + stat_pointinterval(position=position_dodge(.4)) +
    ggh4x::facet_nested_wrap(~Model+bandOrder) + labs(title="learning rate parameter") +
  theme(legend.title = element_blank(), legend.position = "none",plot.title=element_text(hjust=.5))
c_post + lr_post
```

@@

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  left_join(group_ids, by = join_by(id,condit)) |>
  ungroup() |>
  mutate(id = reorder(id, Error_difference)) %>%
  ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
  geom_col() +
  ggh4x::facet_grid2(~condit+bandOrder,axes="all",scales="free_y", independent = "y")+
  labs(fill="Best Model",x="Mean Model Error Difference (ALM - EXAM)",y="Participant")

@@

#### Subjects with biggest differential favoring ALM
```{r}
#| fig-height: 9
#| fig-width: 12
vAlm <- c(245,264,280,408,370,413); cAlm <- c(219,202,225,456,463,407)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit+bandOrder~flab, axes = "all",ncol=3) +
  labs(title="Subjects with biggest differential favoring ALM",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

@@

#### Subjects with biggest differential favoring EXAM
```{r}
#| fig-height: 9
#| fig-width: 12
vAlm <- c(278,215,274,401,375,425); cAlm <- c(260,223,249,376,440,418)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,bandOrder,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit+bandOrder~flab, axes = "all",ncol=3) +
  labs(title="Subjects with biggest differential favoring EXAM",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())
